---
title: "Intro to RMarkdown ADC"
author: "Ellie Honan"
date: "2023-02-27"
output: html_document
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Plain text

whether you *have* any **time** or location constraints (we want to make these experiences as accessible to people as possible so please let us know, for example, if a shorter trip of a couple of days would suit you better than a week trip, or if you ~~have~~ any physical/mobility issues that might restrict where you can visit/the terrain you can work in. Please also indicate how far you would be willing to travel\*)

# Big header

## Small header

### Smaller header

###### Smaller header

...

> test block quote

| Table Header | Second Header |
|--------------|---------------|
| Table Cell   | Cell 2        |
| Cell 3       | Cell 4        |

1.  test ordered list
2.  test item 2
    -   test sub-item 1
    -   test sub-item 2

This is a very interesting fact[@lato2023]

| Table Header | Second Header |
|--------------|---------------|
| Table Cell   | Cell 2        |
| Cell 3       | Cell 4        |

| column 1 | Col2 | Col3 |
|----------|------|------|
| Item 1   |      |      |
| item 2   |      |      |
|          |      |      |

: Example of a table

inline equation: $A = \pi*r^{2}$

summary(penguins)

```{r echo=FALSE}
install.packages("palmerpenguins")
library(palmerpenguins)
summary(penguins)
```

```{r eval=TRUE, echo=FALSE}
paste("Hello", "World!") 
```

```{r}
x <- 4*4

summary(penguins)
```

The shortcut to install a code chunk is ctrl alt what she went so fast













