---
title: "Overview_dplyr"
author: "Ellie Honan"
date: "2023-03-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Libraries

```{r}
library(dplyr)
library(tidyr)
library(readr)
```

## Load Data

```{r}
catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"))
```

The *select* command from the dyplyr package selects **columns**

The *filter* command from the dyplyr package filters **rows**

## Clean Data

```{r}
#r emove the All and notesRegCode columns from the catch_original data frame
catch_data <- catch_original %>% 
  select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum)
# or
# catch_data <- catch_original %>% 
#  select(-All, -notesRegCode)

head(catch_data)
glimpse(catch_data)
```

```{r}
# Mutate Chinook class from character to numeric
catch_clean <- catch_data %>% 
  mutate(Chinook = as.numeric(Chinook))

# search for the na value that gave the warning
i <- which(is.na(catch_clean$Chinook))
# call it
i
#check that line in the original df - it's a I rather than a 1
catch_data[i,]

# change the l to a 1

catch_clean <- catch_data %>% 
  mutate(Chinook = if_else(Chinook == "I", "1", Chinook)) %>% 
  mutate(Chinook = as.integer(Chinook))

glimpse(catch_clean)
```

## Change shape of data using *pivot_longer* and *pivot_wider*

```{r}
# pivot longer practice
catch_long <- catch_clean %>% 
  pivot_longer(cols = -c(Region, Year), names_to = "species", values_to ="catch")
View(catch_long)
# names_to == column names that will go into a new column. In the case above, 
# a new column called "species"

# values_to == values from within those columns, into a new column. 
# In the case above, a new column called "catch"

# pivot wider practice 
catch_wide <- catch_long %>% 
  pivot_wider(names_from = species, values_from = catch)
View(catch_wide)

# so this reverts it back 
```

## Practice *renaming* columns

```{r}
catch_long <- catch_long %>% 
  rename(catch_thousands = catch)

head(catch_long)
```

## Add columns with *mutate* function

```{r}
catch_long <- catch_long %>% 
  mutate(catch = catch_thousands * 1000)
head(catch_long)


# remove catch_thousands
catch_long <- catch_long %>% 
  mutate(catch = catch_thousands * 1000) %>% 
  select(-catch_thousands)

head(catch_long)
```

## Practice using *group_by* and *summarise*

```{r}
mean_region <- catch_long %>% 
  group_by(Region) %>% 
  summarise(catch_mean = mean(catch))

head(mean_region)

# summarise count by region 
n_region <- catch_long %>% 
  group_by(Region) %>%
  summarise(number_of_observations = n(), 
 total_catch = sum(catch)) # number of rows of each group

head(n_region)

```

## Lesson of *filtering*

```{r}
ALU_catch <- catch_long %>% 
  filter(Region == "ALU")

nrow(ALU_catch)

```

## Arrange some rows with *arrange*

```{r}
# this arranges data by default smallest to largest
mean_region_arrange <- catch_long %>% 
  group_by(Region) %>% 
  summarise(mean_catch = mean(catch)) %>% 
  arrange(mean_catch)
# if it was largest to smallest it would be: 
#   arrange(desc(mean_catch))

head(mean_region_arrange)

```

## Practicing *joins*

```{r}
# loading in the byerlyRegionDegs.csv file
# remember this is the mac operator, would need to add url stuff from above
region_defs <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1") %>% 
    select(code, mgmtArea)

head(region_defs)
```

## Join our data

```{r}
catch_joined <- left_join(catch_long, region_defs, by = c("Region" = "code"))

head(catch_joined)

```

```{r}
region_defs <- region_defs %>% 
  rename(Region = code, Region_Name = mgmtArea)

catch_joined <- left_join(catch_long, region_defs, by = "Region")

head(catch_joined)
```

## Practice with *seperate()* and *unite()*

```{r}
sites_df <- data.frame(site = c("HAW-101",
                                "HAW-103",
                                "OAH-320",
                                "OAH-219",
                                "MAI-039"))

sites_df %>%
  separate(site, c("island", "site_number"), "-")

```

```{r}
# Unite things back together
dates_df <- data.frame(year = c("1930",
                                "1930",
                                "1930"),
                       month = c("12",
                                "12",
                                "12"),
                       day = c("14",
                               "15",
                               "16"))

head(dates_df)

dates_df %>% 
  unite(date, year, month, day, sep = "-")
# could also use / in place of - 
```

## Exercise: \*Split the city column in the following data.frame into city and state_code columns\*

```{r}
cities_df <- data.frame(city = c("Juneau AK", 
                                 "Sitka AK", 
                                 "Anchorage AK"))

head(cities_df)

cities_df %>%
  separate(city, c("city", "state code"), " ")

head(cities_df)

```


